TKT RNAseq Experiment: Differential Expression Analysis

### 2023-06-20 TKT_RNAseq/R Code/EdgeR_Analysis.Rmd
Update with refined normalization

---
title: "TKT RNAseq Experiment: Differential Expression Analysis"
author: "John Santiago"
date: "2022-03-23"
output: 
  bookdown::html_document2:
    number_sections: false
---

@@

- #### Figure \@ref(fig:fig1): MDS plot

@@

### TKT effect in female G85R
- #### Figure \@ref(fig:fig3): Heatmaps of Total DEGs in GR.FC x GR.FDf

- #### Figure \@ref(fig:fig4): Heatmaps of Total DEGs in GR.FC x GR.FOE

### TKT effect in female WT
- #### Figure \@ref(fig:fig5): Heatmaps of Total DEGs in WT.FC x WT.FDf

- #### Figure \@ref(fig:fig6): Heatmaps of Total DEGs in WT.FC x WT.FOE

### Genotype effect between female G85R and WT
- #### Figure \@ref(fig:fig7): Heatmaps of Total DEGs in GR.FC x WT.FC

- #### Figure \@ref(fig:fig8): Heatmaps of Total DEGs in GR.FDf x WT.FDf

- #### Figure \@ref(fig:fig9): Heatmaps of Total DEGs in GR.FOE x WT.FOE

### TKT effect in Male G85R
- #### Figure \@ref(fig:fig10): Heatmaps of Total DEGs in GR.MC x GR.MDf

- #### Figure \@ref(fig:fig11): Heatmaps of Total DEGs in GR.MC x GR.MOE

### TKT effect in Male WT
- #### Figure \@ref(fig:fig12): Heatmaps of Total DEGs in WT.MC x WT.MDf

- #### Figure \@ref(fig:fig13): Heatmaps of Total DEGs in WT.MC x WT.MOE

### Genotype effect between Male G85R and WT
- #### Figure \@ref(fig:fig14): Heatmaps of Total DEGs in GR.MC x WT.MC

- #### Figure \@ref(fig:fig15): Heatmaps of Total DEGs in GR.MDf x WT.MDf

- #### Figure \@ref(fig:fig16): Heatmaps of Total DEGs in GR.MOE x WT.MOE

@@

MDSdata = plotMDS(z, col=as.numeric(factor(gr)))

temp=convert[,"Symbol"]
temp2=setdiff(row.names(cpmdata),convert[,2])
names(temp)=convert[,2]
names(temp2)=temp2
convert=c(temp,temp2)

@@

```{r fig1, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot.'}

@@

```{r fig2, echo = F, fig.cap='Barplot of Total DEGs.'}

@@

```{r echo = F}

@@

```{r fig3, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.FC x GR.FDf.'}

Samples = row.names(groups)[groups$Sex =='F'
                            & groups$Genotype == 'GR']

hmdata = cpmdata[Comparisons[[1]],Samples]
##max.hmdata = apply(hmdata, 1, max)
##hmdata = hmdata[max.hmdata > 5,]

@@

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "GR.FC x GR.FDf",
            custom_hovertext = cpm.text)

@@

```{r fig4, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.FC x GR.FOE.'}

Samples = row.names(groups)[groups$Sex =='F'
                            & groups$Genotype == 'GR']

hmdata = cpmdata[Comparisons[[2]],Samples]

##max.hmdata = apply(hmdata, 1, max)
##hmdata = hmdata[max.hmdata > 5,]

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "GR.FC x GR.FOE",
            custom_hovertext = cpm.text)

```{r fig5, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in WT.FC x WT.FDf.'}

Samples = row.names(groups)[groups$Sex =='F'
                            & groups$Genotype == 'WT']

hmdata = cpmdata[Comparisons[[3]],Samples]

##max.hmdata = apply(hmdata, 1, max)
##hmdata = hmdata[max.hmdata > 5,]

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "WT.FC x WT.FDf",
            custom_hovertext = cpm.text)

```{r fig6, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in WT.FC x WT.FOE.'}

Samples = row.names(groups)[groups$Sex =='F'
                            & groups$Genotype == 'WT']

hmdata = cpmdata[Comparisons[[4]],Samples]

##max.hmdata = apply(hmdata, 1, max)
##hmdata = hmdata[max.hmdata > 5,]

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "WT.FC x WT.FOE",
            custom_hovertext = cpm.text)

```{r fig7, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.FC x WT.FC.'}

Samples = row.names(groups)[groups$Sex =='F'
                            & groups$TKT == 'Control']

hmdata = cpmdata[Comparisons[[5]],Samples]

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "GR.FC x WT.FC",
            custom_hovertext = cpm.text)

```{r fig8, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.FDf x WT.FDf.'}

Samples = row.names(groups)[groups$Sex =='F'
                            & groups$TKT == 'DF']

hmdata = cpmdata[Comparisons[[6]],Samples]

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "GR.FDf x WT.FDf",
            custom_hovertext = cpm.text)

```{r fig9, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.FOE x WT.FOE.'}

Samples = row.names(groups)[groups$Sex =='F'
                            & groups$TKT == 'OE']

hmdata = cpmdata[Comparisons[[7]],Samples]

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "GR.FOE x WT.FOE",
            custom_hovertext = cpm.text)

```{r fig10, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.MC x GR.MDf.'}

Samples = row.names(groups)[groups$Sex =='M'
                            & groups$Genotype == 'GR']

hmdata = cpmdata[Comparisons[[8]],Samples]
##max.hmdata = apply(hmdata, 1, max)
##hmdata = hmdata[max.hmdata > 5,]

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "GR.MC x GR.MDf",
            custom_hovertext = cpm.text)

```{r fig11, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.MC x GR.MOE.'}

Samples = row.names(groups)[groups$Sex =='M'
                            & groups$Genotype == 'GR']

hmdata = cpmdata[Comparisons[[9]],Samples]

##max.hmdata = apply(hmdata, 1, max)
##hmdata = hmdata[max.hmdata > 5,]

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = row.names(hmdata),
            key = T,
            column_text_angle = 90,
            col_side_colors=groups[Samples,1:3],
            col_side_palette=Spectral,
            main= "GR.MC x GR.MOE",
            custom_hovertext = cpm.text)

```{r fig12, echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in WT.MC x WT.MDf.'}

Samples = row.names(groups)[groups$Sex =='M'
                            & groups$Genotype == 'WT']

hmdata = cpmdata[Comparisons[[10]],Samples]

##max.hmdata = apply(hmdata, 1, max)
##hmdata = hmdata[max.hmdata > 5,]

>>>>

---
title: "TKT RNAseq Experiment: Differential Expression Analysis"
author: "John Santiago"
date: "2022-06-20"
output: 
  bookdown::html_document2:
    number_sections: false
---

@@

### All samples processed together
- #### Figure \@ref(fig:fig1): MDS plot, all samples

@@

### Male and female samples processed together with TEs and outlier libraries removed
- #### Figure \@ref(fig:fig3): MDS plot, all samples

- #### Figure \@ref(fig:fig4): Barplot of Total DEGs

### Male and Female samples processed separately with TEs and outlier libraries removed
- #### Figure \@ref(fig:fig5): MDS plot of female samples only

- #### Figure \@ref(fig:fig6): MDS plot of male samples only

- #### Figure \@ref(fig:fig7): Barplot of Total DEGs

@@

MDSdata = plotMDS(z, top = 13582, col=as.numeric(factor(gr)))

temp=convert[,"Symbol"]
temp2=setdiff(row.names(countdata),convert[,2])
names(temp)=convert[,2]
names(temp2)=temp2
convert=c(temp,temp2)

@@

```{r fig1, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot, all samples.'}

@@

```{r fig2, echo = F, fig.cap='Barplot of Total DEGs, all samples.'}

@@

```{r, include = F}
###EdgeR with TEs and outlier libraries removed

groups2 = groups[setdiff(row.names(groups), c("GR.F3", "TktDfWT.F3")),]##"WT.F3",

countdata2 = countdata[, row.names(groups2)]

countdata2 = countdata2[-(grep("ti",row.names(countdata2))),]
countdata2 = countdata2[-(grep("trans",row.names(countdata2))),]

x <- countdata2
gr <- factor(groups2$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata2=cpm(z)

##GR.F TKT control vs TKT altered
compare = makeContrasts(TktDfGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxDf2=DEout
sGR.F.CxDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxOE2=DEout
sGR.F.CxOE2=DEout[DEout$FDR<.05,]

##WT.F TKT control vs TKT altered
compare = makeContrasts(TktDfWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxDf2=DEout
sWT.F.CxDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxOE2=DEout
sWT.F.CxOE2=DEout[DEout$FDR<.05,]

##WT.F vs GR.F
compare = makeContrasts(GR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.F2=DEout
sGRxWT.F2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.F-TktDfWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FDf2=DEout
sGRxWT.FDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-TktOEWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FOE2=DEout
sGRxWT.FOE2=DEout[DEout$FDR<.05,]

##GR.M TKT control vs TKT altered
compare = makeContrasts(TktDfGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxDf2=DEout
sGR.M.CxDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxOE2=DEout
sGR.M.CxOE2=DEout[DEout$FDR<.05,]

##WT.M TKT control vs TKT altered
compare = makeContrasts(TktDfWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxDf2=DEout
sWT.M.CxDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxOE2=DEout
sWT.M.CxOE2=DEout[DEout$FDR<.05,]

##WT.M vs GR.M
compare = makeContrasts(GR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.M2=DEout
sGRxWT.M2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-TktDfWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MDf2=DEout
sGRxWT.MDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-TktOEWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MOE2=DEout
sGRxWT.MOE2=DEout[DEout$FDR<.05,]

MDSdata2 = plotMDS(z, top = 13086, col=as.numeric(factor(gr)))

Comparisons2 = list(row.names(sGR.F.CxDf2), row.names(sGR.F.CxOE2),
                row.names(sWT.F.CxDf2), row.names(sWT.F.CxOE2),
                row.names(sGRxWT.F2), row.names(sGRxWT.FDf2), row.names(sGRxWT.FOE2),
                row.names(sGR.M.CxDf2), row.names(sGR.M.CxOE2),
                row.names(sWT.M.CxDf2), row.names(sWT.M.CxOE2),
                row.names(sGRxWT.M2), row.names(sGRxWT.MDf2), row.names(sGRxWT.MOE2))

```{r fig3, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot, all samples, TEs and outlier libraries removed.'}

MDSplot = data.frame(x = MDSdata2$x,
                     y = MDSdata2$y,
                     sample = row.names(groups2))
MDSplot = cbind(MDSplot, groups[MDSplot$sample,])
MDSplot$GenoSex = paste0(MDSplot$Genotype,MDSplot$Sex)
MDSplot = MDSplot[order(MDSplot$TKT),]
MDSplot = MDSplot[order(MDSplot$Sex),]
MDSplot = MDSplot[order(MDSplot$GenoSex),]
MDSplot$Group = factor(MDSplot$Group, levels = unique(MDSplot$Group))
##MDSplot$GenoSex = factor(MDSplot$GenoSex, levels = unique(MDSplot$GenoSex))

  fig  = plot_ly(data = MDSplot, 
                 x = ~x, 
                 y = ~y,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Group,
                 ##legendrank = ~ as.numeric(Group),
                 symbols = rep(c('circle', 'triangle-down', 'triangle-up'),4),
                 colors = c("red","firebrick3","deeppink",
                            "lightskyblue","royalblue","cyan",
                            "gold", "darkgoldenrod1", "yellow",
                            "limegreen", "darkgreen", "chartreuse"),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", MDSplot$Sample)) %>%
    layout(xaxis = list(title = paste0("Leading logFC dim 1 ", "(", round((MDSdata2$var.explained[1]*100),1), "%)")),
           yaxis = list(title = paste0("Leading logFC dim 2 ", "(", round((MDSdata2$var.explained[2]*100),1), "%)")))

@@

```{r fig4, echo = F, fig.cap='Barplot of Total DEGs, all samples, TEs and outlier libraries removed.'}

Female = c(nrow(sGR.F.CxDf2), nrow(sGR.F.CxOE2),
           nrow(sWT.F.CxDf2), nrow(sWT.F.CxOE2),
           nrow(sGRxWT.F2), nrow(sGRxWT.FDf2), nrow(sGRxWT.FOE2))

Male = c(nrow(sGR.M.CxDf2), nrow(sGR.M.CxOE2),
         nrow(sWT.M.CxDf2), nrow(sWT.M.CxOE2),
         nrow(sGRxWT.M2), nrow(sGRxWT.MDf2), nrow(sGRxWT.MOE2))

```{r, include = F}
###EdgeR with TEs and outlier libraries removed, Females only

groups3 = groups[setdiff(row.names(groups), c("GR.F3", "WT.F3","TktDfWT.F3")),]##

groups3 = groups3[groups3$Sex == "F",]

countdata3 = countdata[, row.names(groups3)]

countdata3 = countdata3[-(grep("ti",row.names(countdata3))),]
countdata3 = countdata3[-(grep("trans",row.names(countdata3))),]

x <- countdata3
gr <- factor(groups3$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata3=cpm(z)

##GR.F TKT control vs TKT altered
compare = makeContrasts(TktDfGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxDf3=DEout
sGR.F.CxDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxOE3=DEout
sGR.F.CxOE3=DEout[DEout$FDR<.05,]

##WT.F TKT control vs TKT altered
compare = makeContrasts(TktDfWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxDf3=DEout
sWT.F.CxDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxOE3=DEout
sWT.F.CxOE3=DEout[DEout$FDR<.05,]

##WT.F vs GR.F
compare = makeContrasts(GR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.F3=DEout
sGRxWT.F3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.F-TktDfWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FDf3=DEout
sGRxWT.FDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-TktOEWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FOE3=DEout
sGRxWT.FOE3=DEout[DEout$FDR<.05,]

MDSdata3 = plotMDS(z, top = 10968, col=as.numeric(factor(gr)))

```{r fig5, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot, females only, TEs and outliers removed.'}

MDSplot = data.frame(x = MDSdata3$x,
                     y = MDSdata3$y,
                     sample = row.names(groups3))
MDSplot = cbind(MDSplot, groups[MDSplot$sample,])
MDSplot$GenoSex = paste0(MDSplot$Genotype,MDSplot$Sex)
MDSplot = MDSplot[order(MDSplot$TKT),]
MDSplot = MDSplot[order(MDSplot$Sex),]
MDSplot = MDSplot[order(MDSplot$GenoSex),]
MDSplot$Group = factor(MDSplot$Group, levels = unique(MDSplot$Group))
##MDSplot$GenoSex = factor(MDSplot$GenoSex, levels = unique(MDSplot$GenoSex))

  fig  = plot_ly(data = MDSplot, 
                 x = ~x, 
                 y = ~y,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Group,
                 ##legendrank = ~ as.numeric(Group),
                 symbols = rep(c('circle', 'triangle-down', 'triangle-up'),4),
                 colors = c("red","firebrick3","deeppink",
                            "gold", "darkgoldenrod1", "yellow"),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", MDSplot$Sample)) %>%
    layout(xaxis = list(title = paste0("Leading logFC dim 1 ", "(", round((MDSdata3$var.explained[1]*100),1), "%)")),
           yaxis = list(title = paste0("Leading logFC dim 2 ", "(", round((MDSdata3$var.explained[2]*100),1), "%)")))

```{r, include = F}
###EdgeR with TEs and outlier libraries removed, Males only

groups4 = groups[groups$Sex == "M",]

countdata4 = countdata[, row.names(groups4)]

countdata4 = countdata4[-(grep("ti",row.names(countdata4))),]
countdata4 = countdata4[-(grep("trans",row.names(countdata4))),]

x <- countdata4
gr <- factor(groups4$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata4=cpm(z)

##GR.M TKT control vs TKT altered
compare = makeContrasts(TktDfGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxDf3=DEout
sGR.M.CxDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxOE3=DEout
sGR.M.CxOE3=DEout[DEout$FDR<.05,]

##WT.F TKT control vs TKT altered
compare = makeContrasts(TktDfWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxDf3=DEout
sWT.M.CxDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxOE3=DEout
sWT.M.CxOE3=DEout[DEout$FDR<.05,]

##WT.F vs GR.F
compare = makeContrasts(GR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.M3=DEout
sGRxWT.M3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-TktDfWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MDf3=DEout
sGRxWT.MDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-TktOEWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MOE3=DEout
sGRxWT.MOE3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRMDfxWTM=DEout
sGRMDfxWTM=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-TktDfWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRMDfxWTMDf=DEout
sGRMDfxWTMDf=DEout[DEout$FDR<.05,]

MDSdata4 = plotMDS(z, top = 12764, col=as.numeric(factor(gr)))

```{r fig6, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot, males, TEs and outlier libraries removed.'}

MDSplot = data.frame(x = MDSdata4$x,
                     y = MDSdata4$y,
                     sample = row.names(groups4))
MDSplot = cbind(MDSplot, groups[MDSplot$sample,])
MDSplot$GenoSex = paste0(MDSplot$Genotype,MDSplot$Sex)
MDSplot = MDSplot[order(MDSplot$TKT),]
MDSplot = MDSplot[order(MDSplot$Sex),]
MDSplot = MDSplot[order(MDSplot$GenoSex),]
MDSplot$Group = factor(MDSplot$Group, levels = unique(MDSplot$Group))
##MDSplot$GenoSex = factor(MDSplot$GenoSex, levels = unique(MDSplot$GenoSex))

  fig  = plot_ly(data = MDSplot, 
                 x = ~x, 
                 y = ~y,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Group,
                 ##legendrank = ~ as.numeric(Group),
                 symbols = rep(c('circle', 'triangle-down', 'triangle-up'),4),
                 colors = c("lightskyblue","royalblue","cyan",
                            "limegreen", "darkgreen", "chartreuse"),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", MDSplot$Sample)) %>%
    layout(xaxis = list(title = paste0("Leading logFC dim 1 ", "(", round((MDSdata4$var.explained[1]*100),1), "%)")),
           yaxis = list(title = paste0("Leading logFC dim 2 ", "(", round((MDSdata4$var.explained[2]*100),1), "%)")))

```{r fig7, echo = F, fig.cap='Barplot of Total DEGs, males and females processed separately, TEs and outlier libraries removed.'}

Female = c(nrow(sGR.F.CxDf3), nrow(sGR.F.CxOE3),
           nrow(sWT.F.CxDf3), nrow(sWT.F.CxOE3),
           nrow(sGRxWT.F3), nrow(sGRxWT.FDf3), nrow(sGRxWT.FOE3))

Male = c(nrow(sGR.M.CxDf3), nrow(sGR.M.CxOE3),
         nrow(sWT.M.CxDf3), nrow(sWT.M.CxOE3),
         nrow(sGRxWT.M3), nrow(sGRxWT.MDf3), nrow(sGRxWT.MOE3))

```{r echo = F, fig.cap='Barplot of Total DEGs excluding TEs.'}

Female.trim = c(nrow(sGR.F.CxDf2), nrow(sGR.F.CxOE2),
           nrow(sWT.F.CxDf2), nrow(sWT.F.CxOE2),
           nrow(sGRxWT.F2), nrow(sGRxWT.FDf2), nrow(sGRxWT.FOE2))

Male.trim = c(nrow(sGR.M.CxDf2), nrow(sGR.M.CxOE2),
         nrow(sWT.M.CxDf2), nrow(sWT.M.CxOE2),
         nrow(sGRxWT.M2), nrow(sGRxWT.MDf2), nrow(sGRxWT.MOE2))

data = data.frame(Female, Female.trim, Male, Male.trim, Comparison)

fig <- fig %>% add_trace(y = ~Female.trim, name = 'Trim Female',
                         marker = list(color = c("pink"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Male.trim, name = 'Trim Males',
                         marker = list(color = c("skyblue"),
                                       line = list(color = "black", width = 1.5)))

```{r echo = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.FC x GR.FDf.'}

Samples = row.names(groups2)[groups2$Sex =='M'
                            & groups2$Genotype == 'GR']

hmdata = cpmdata2[Comparisons2[[8]],Samples]

hmdata=hmdata[2500:3000,]

hmdata=hmdata[intersect(row.names(hmdata),row.names(cpmdata)),]

hmdata=cbind(hmdata, cpmdata[row.names(hmdata),colnames(hmdata)])

colnames(hmdata)[10:18] = paste0("Old.",colnames(hmdata)[10:18])

@@

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = as.character(row.names(hmdata)),
            key = T,
            column_text_angle = 90,
            ##col_side_colors=groups[Samples,1:3],
            ##col_side_palette=Spectral,
            main= "GR.FC x GR.FDf",
            custom_hovertext = cpm.text)

@@
